Preparing the NHANES Data for the Replication of: Association between sleep duration on workdays and blood pressure in non‑overweight/obese population in NHANES: a public database research

### 2023-12-09 manuscript/paper.Rmd
upd

```{r source-file, echo=FALSE, include=FALSE}
# source file 
source(file = "/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/ProjectFinal/data/scrip2.R")
```
# Introduction

@@

Our current analyses combine the 2015 - 2018 NHANES survey cycles to yield $n = 19225$ observations on 35 covariates. This sample size was comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and 2017/2018 survey cycles, respectively. Analyses excluded individuals with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055), and body mass index (BMI) or those with BMI > 25 kg/$m^2$ (overweight) (n = 5521). Individuals that reported being on anti-hypertensive medication were also excluded from further analyses (n = 2944).

@@

Sleep duration on workdays was evaluated by the questionnaire with the
following questions: "Number of hours usually sleep on weekdays or
workdays". Sleep duration was divided into three groups, which were < 6
h, 6--8 h, ≥ 8 h respectively, of which 6--8 h was used as the reference
group.

@@

In addition to the exposure and outcome variables, our analyses included the following as covariates: Race divided into four groups as Mexican American, white, black and other race. Alcohol consumption was grouped into drinking, no drinking, not recorded. Smoking status as smoking, not smoking, not recorded. Diabetes was defined as yes, no, borderline, or not recorded. Hypertension was defined as yes, no, or not
recorded. Snorting was defined as yes, no, and not recorded. US citizenship status was defined as citizen by birth or naturalization, don't know, not a citizen, or refused to answer. Education level was grouped into four categories including graduate studies, high school, less than grade 12 or some college. Additional covariates included marital status, gender, age, albumin, creatinine,hemoglobin, total cholesterol (TC), aspertate aminotransferase (AST), high-density lipoprotein (HDL), and body mass index (BMI). Details about these variables can be found at https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx. Age, albumin, creatinine, hemoglobin, TC, AST, HDL, and BMI were analyzed as continuous variables whereas gender, alcohol consumption, diabetes, smoking,
race, hypertension, and snorting were analyzed as categorical variables. Table \@ref{tab:tabone} highlights the variable names and descriptions as utilized in this study.

```{r tabone, echo=FALSE}
tab.one <- kable(vars, caption = "Variable names and descriptions", 
                 format = "latex") %>% 
  kable_styling(latex_options = c("HOLD_position")) %>% 
  row_spec(c(0), bold = TRUE) %>% 
  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>% 
  column_spec(3, width = "8cm")
tab.one
```

@@

@@

Table \@ref(tab:tabtwo) summarizes participant characteristics in the 2015 - 2018 NHANES survey cycles stratified by gender. Among the participants, the proportion of males and females were 44.81% (n = 1378) and 55.19% (n = 1509),
respectively. With ethnicity, the proportion of Mexican American, White and Black were 6.27%, 64.62% and
9.94%, respectively. Overall, the mean (SD) values for age, albumin, creatinine, SBP, DBP, hemoglobin, TC, AST,
HDL, BMI were 38.54 (17.72) years, 43.48 (3.58) g/L, 73.21 (19.73) umol/L, 115.33 (14.56) mmHg, 68.47 (10.08)
mmHg, 14.07(1.44)g/dL, 4.67 (1.01) mmol/L, 23.18 (13.01) IU/L, 1.59 (0.43) mmol/L, and 21.97(2.06) kg/m2,
respectively. Among the participants, 66.01% were consumed alcohol drinkers, 2.61% were diabetes, 3.16% were
hypertension, 6.72% have snort or stop breathing. 17.99% were smokers. Sleep duration was divided into three
groups, which were < 6 h, 6–8 h, ≥ 8 h, each with a proportion of 6.26%, 40.53%, 53.21%, respectively.

@@

```{r tabtwo, echo=FALSE}
tabsum %>% 
  as_kable_extra(format = "latex", caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", digits = 2, longtable = T) %>% 
  kable_styling(latex_options = c("HOLD_position"))
```

>>>>

```{r source-file, echo=FALSE, include=FALSE}
# source file 
source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
```

# Introduction

@@

Our current analyses combine the 2015 - 2018 NHANES survey cycles to
yield $n = 19225$ observations on 35 covariates. This sample size was
comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and
2017/2018 survey cycles, respectively. Analyses excluded individuals
with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055),
and body mass index (BMI) or those with BMI \> 25 kg/$m^2$ (overweight)
(n = 5521). Individuals that reported being on anti-hypertensive
medication were also excluded from further analyses (n = 2944).

@@

Sleep duration on workdays was evaluated by the questionnaire with the
following questions: "Number of hours usually sleep on weekdays or
workdays". Sleep duration was divided into three groups, which were \< 6
h, 6--8 h, ≥ 8 h respectively, of which 6--8 h was used as the reference
group.

@@

In addition to the exposure and outcome variables, our analyses included
the following as covariates: Race divided into four groups as Mexican
American, white, black and other race. Alcohol consumption was grouped
into drinking, no drinking, not recorded. Smoking status as smoking, not
smoking, not recorded. Diabetes was defined as yes, no, borderline, or
not recorded. Hypertension was defined as yes, no, or not recorded.
Snorting was defined as yes, no, and not recorded. US citizenship status
was defined as citizen by birth or naturalization, don't know, not a
citizen, or refused to answer. Education level was grouped into four
categories including graduate studies, high school, less than grade 12
or some college. Additional covariates included marital status, gender,
age, albumin, creatinine,hemoglobin, total cholesterol (TC), aspertate
aminotransferase (AST), high-density lipoprotein (HDL), and body mass
index (BMI). Details about these variables can be found at
<https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx>. Age, albumin,
creatinine, hemoglobin, TC, AST, HDL, and BMI were analyzed as
continuous variables whereas gender, alcohol consumption, diabetes,
smoking, race, hypertension, and snorting were analyzed as categorical
variables. Table \@ref{tab:tabone} highlights the variable names and
descriptions as utilized in this study.

```{r tabone, echo=FALSE}
#tab.one <- as_kable_extra(vars, caption = "Variable names and descriptions", 
#                 format = "latex", booktabs = TRUE) %>% 
#  kable_styling(latex_options = c("HOLD_position") %>% 
#  row_spec(c(0), bold = TRUE) %>% 
#  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>% 
#  column_spec(3, width = "8cm")
#tab.one

vars %>% kable(booktabs = T) %>% kable_styling(latex_options = "scale_down")

```

@@

In order to select variables impacting blood pressure, we proceed by
doing some variable selection with regularization techniques. We first
employ best subset selection by using 10-fold cross validation to select
the best number of subsets. For syastolic blood pressure, forward best
subset selection choose 13 variables and for diastolic blood pressure 19
variables. These variables are showed, along with its significance in
Table \@ref(tab:tabreg)

```{r tabref, echo=FALSE}
tab.ref %>% as_kable_extra(booktabs = T, caption ="Coefficients for Foward Best Subset Selection") %>% kable_styling(latex_options = "scale_down")

```

@@

Table \@ref(tab:tabtwo) summarizes participant characteristics in the
2015 - 2018 NHANES survey cycles stratified by gender. Among the
participants, the proportion of males and females were 44.81% (n = 1378)
and 55.19% (n = 1509), respectively. With ethnicity, the proportion of
Mexican American, White and Black were 6.27%, 64.62% and 9.94%,
respectively. Overall, the mean (SD) values for age, albumin,
creatinine, SBP, DBP, hemoglobin, TC, AST, HDL, BMI were 38.54 (17.72)
years, 43.48 (3.58) g/L, 73.21 (19.73) umol/L, 115.33 (14.56) mmHg,
68.47 (10.08) mmHg, 14.07(1.44)g/dL, 4.67 (1.01) mmol/L, 23.18 (13.01)
IU/L, 1.59 (0.43) mmol/L, and 21.97(2.06) kg/m2, respectively. Among the
participants, 66.01% were consumed alcohol drinkers, 2.61% were
diabetes, 3.16% were hypertension, 6.72% have snort or stop breathing.
17.99% were smokers. Sleep duration was divided into three groups, which
were \< 6 h, 6--8 h, ≥ 8 h, each with a proportion of 6.26%, 40.53%,
53.21%, respectively.

@@

```{r tabtwo, echo=FALSE}
#tabsum %>% 
#  as_kable_extra(format = "latex", caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", digits = 2, longtable = T) %>% 
#  kable_styling(latex_options = c("HOLD_position"))

tabsum %>%
  as_kable_extra(booktabs = TRUE, 
                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", 
                 longtable = TRUE) %>%
  kable_styling(latex_options = "scale_down")
```
